---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

    
# scennep: Single-Cell Nearest-Neighbor Pseudobulking

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- [![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable) -->
<!-- badges: end -->

`scennep` computes pseudobulk expression profiles for each cell in a single-cell RNA-seq (scRNA-seq) dataset based on its nearest neighbors. This approach aggregates the expression values of a cell and its top nearest neighbors.
The purpose is to rescue zeros caused by dropouts and transcriptional bursting.
`scennep` has many use cases - it was developed to improve the integration of scRNA-seq and cytometry data.

## Article

(TODO) A preprint is soon available for `scennep`.

Please cite with `citation("scennep")`

## Installation

### Dependencies

`scennep` works with both `Bioconductor` objects and `Seurat`.
Install your preferred framework or both:

```{r dependencies, eval=FALSE}
# Seurat
install.packages("Seurat")

# Bioconductor
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(c("SingleCellExperiment", "scater", "scuttle", "scran", "igraph"))
```

### Install from GitHub

```{r install, eval=FALSE}
remotes::install_github("shdam/scennep")
```

## Vignettes

(TODO) View use case vignettes on [Biosurf](https://biosurf.org/).

## Usage

`scennep` will work on a `matrix`, `Seurat`, and `SingleCellExperiment` object.
If a matrix is provided, please choose an option for `as = c("seurat", "bioc")`.

This example usage uses a tiny simulated dataset as a `Seurat` object.

    
```{r scennep}
library(Seurat)
library(scennep)
data("pbmc_small")

# Current PCA plot
Seurat::DimPlot(pbmc_small, reduction = "pca", group.by = "RNA_snn_res.1")

# Compute pseudobulk expression profiles
pbmc_small <- scennep(
    pbmc_small,
    nn_count = 5 # Number of nearest neighbors - default: 20 - 5 is just for this tiny example
    )

```

```{r plotting}

# PCA of scennep output
pbmc_small <- Seurat::ScaleData(pbmc_small)
pbmc_small <- Seurat::RunPCA(pbmc_small, verbose = FALSE)
Seurat::DimPlot(pbmc_small, reduction = "pca", group.by = "RNA_snn_res.1")
```


## Report issues

If you have any issues or questions regarding the use of `scennep`, 
please do not hesitate to raise an issue on GitHub. 
In this way, others may also benefit from the answers and discussions.